---
title: "Mini-Project 02"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 02


```{r}
library(dplyr)
library(ggplot2)

# Read the dataset
data <- read.csv("../data/all_billboard_summer_hits.csv")

```

```{r}
summary(data)
data$year <- as.Date(data$year)

```
```{r}
ggplot(data, aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Danceability")

```
```{r}

ggplot(data, aes(x = energy)) +
  geom_histogram(binwidth = 0.05, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Energy")

```

```{r}
correlation_matrix <- cor(data[, c("danceability", "energy", "loudness", "tempo")])
print(correlation_matrix)
```
```{r}
trends_over_time <- data %>%
  group_by(year) %>%
  summarise(avg_danceability = mean(danceability),
            avg_energy = mean(energy))

ggplot(trends_over_time, aes(x = year, y = avg_danceability)) +
  geom_line(color = "blue", size = 1.2) +
  geom_line(aes(y = avg_energy), color = "red", size = 1.2) +
  labs(title = "Trends in Danceability and Energy Over Time", x = "Year", y = "Average Value") +
  theme_minimal()

```


```{r}
artist_track_distribution <- data %>%
  group_by(artist_name) %>%
  summarise(total_tracks = n()) %>%
  arrange(desc(total_tracks)) %>%
  top_n(10)

ggplot(artist_track_distribution, aes(x = reorder(artist_name, total_tracks), y = total_tracks)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  coord_flip() +
  labs(title = "Number of Tracks per Artist (Top 10)", x = "Artist Name", y = "Number of Tracks")

```

```{r}
library(plotly)
library(dplyr)

data$year <- as.Date(data$year)

trends_over_time <- data %>%
  group_by(year) %>%
  summarise(avg_instrumentalness = mean(instrumentalness),
            avg_speechiness = mean(speechiness))

interactive_plot <- plot_ly(trends_over_time, x = ~year, y = ~avg_instrumentalness, name = 'Instrumentalness', type = 'scatter', mode = 'lines') %>%
  add_trace(y = ~avg_speechiness, name = 'Speechiness') %>%
  layout(title = 'Trends in Instrumentalness and Speechiness Over Time',
         xaxis = list(title = 'Year'),
         yaxis = list(title = 'Average Value'),
         legend = list(title = 'Feature'))

htmlwidgets::saveWidget(as_widget(interactive_plot), "interactive_plot_instrumentalness_speechiness.html")

```

```{r}

library(GGally)

numeric_data <- select(data, danceability, energy, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo)

pair_plot <- ggpairs(numeric_data)

print(pair_plot)

```
```{r}
library(ggplot2)
library(sf)

shapefile <- st_read("../data/Florida_Counties.shp")

ggplot() +
  geom_sf(data = shapefile)
```
## Introduction

This is my submission for Data Visualization and Reproducible Research Mini Project 2.

## Data Overview

The data set I used was [all_billboard_summer_hits.csv](https://raw.githubusercontent.com/reisanar/datasets/master/all_billboard_summer_hits.csv)

## Data Cleaning and Preparation

The data did not require much processing however I did format the date to a date for easier analysis.


## Initial Plans
I planned to make a chart that showed 2 trend lines in order to see how they diverge or converge. With music, there are trends and seasonal interests so i expected that to show in the data.

## The story
Rihanna is incredibly dominant however this may not be surprising. I was, however, surprised to see names I had never heard of before such as "Wings" and "Donna Summer". They were incredibly high up considering they were foreign names to me.

## Principles
I Did my best to keep my graphs clean and readable. I focused on them being lightweight, not visually noisy, and efficiently communicating information.